Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-11-10 prop_letter.tex
prop_letter.tex \newcommand{\eventsC}{\ensuremath{\events_{C}}} Defined to satisfy the different requirement for different journals...

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\gerardo}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{OliveGreen}{Gerardo says: #1}\end{minipage}}}
\newcommand{\andy}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{blue}{Andy says: #1}\end{minipage}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\rmi}{\mathrm{i}}
\lineskiplimit=1pt
\newcommand{\ultramodular}{\mathcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{\mathrm{L{#1}}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{\mathrm{R{#1}}}}}
\newcommand{\muB}{\ensuremath{\mu^{B}}}

@@

We now have the necessary ingredients to define the quantum extension,
QIVPMs, as a generalization of both classical IVPMs and conventional
quantum probability measures. We will show that QIVPMs reduce to
classical IVPMs when the space of quantum events $\events$ is
restricted to mutually commuting events $\events_{\mathrm{C}}$, i.e.,
to compatible events that can be measured simultaneously. In
Sec.~\ref{sec:Gleason} we will discuss the connection between QIVPMs
and conventional quantum probability measures in detail.

@@

The proof that QIVPMs become classical IVPMs when the space of events
includes only mutually commuting events is presented in detail in the
long version of this paper~\cite{HOSTunpublished}. The idea of the
proof is as follows. Given a subspace of mutually commuting
events~$\events_{\mathrm{C}}$, we can construct a partial orthonormal
basis by diagonalization and complete it to a full orthonormal basis
$\overline{\events_{\mathrm{C}}}$. We then build a bijection between
the QIVPM on the set of projectors associated with this basis and the
set of classical events corresponding to this basis. This
correspondence leads to a crucial property in our setting.

\begin{thm}[Non-empty Core for Compatible Measurements] \label{thm:Shapley}
  For QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace
  of events~$\events_{\mathrm{C}}\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\events_{\mathrm{C}}\right)\ne\emptyset$.
\end{thm}

This theorem follows directly from our
correspondence~\cite{HOSTunpublished} and Shapley's
theorem~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016},
which proves that a classical IVPM always contains at least one state $\rho$
that is consistent with every event in $\events_{\mathrm{C}}$. In contrast, as we
illustrate in Sec.~\ref{sec:Gleason}, the set of states
$\coreBorn\left(\bar{\mu},\events'\right)$ that is consistent with
some arbitrary (not necessarily commuting) events $\events'$ and a
particular QIVPM $\bar{\mu}$ may be \emph{empty}.

@@

\begin{thm}[Equivalence of Ultramodular Functions and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}A function~$\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$
is ultramodular.
\item \label{enu:iota-mu-CIVPM}The composite
  function~$\ultramodular\circ\mu:\events_{\mathrm{C}}\rightarrow\mathscr{I}$
  is a classical IVPM for all classical probability
  measure~$\mu:\events_{\mathrm{C}}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composite
  function~$\ultramodular\circ\mu:\events\rightarrow\mathscr{I}$ is a
  QIVPM for all quantum probability
  measure~$\mu:\events\rightarrow\left[0,1\right]$.
\end{enumerate}
\end{thm}

Statement~\ref{enu:iota-subject-to} implies \ref{enu:iota-mu-CIVPM}
and~\ref{enu:iota-mu-QIVPM} as we have outlined above. Conversely, for the quantum case, we want
to show that if $\ultramodular$ is not ultramodular, then for
some quantum probability measure $\mu$, the composite
$\ultramodular\circ\mu$ might not be a QIVPM. Suppose there are three
particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $y=x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but they don't
satisfy Eq.~(\ref{eq:iota-Inclusion}). Consider the state:
\[
\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-y\right)\proj{3}\,.
\]
The induced map using the Born rule and blurred by $\ultramodular$,
$\ultramodular\circ\muB_{\rho}$, fails to satisfy
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}) when
$P_{0}=\proj{0}+\proj{2}$ and $P_{1}=\proj{1}+\proj{2}$. In other
words, this induced map fails to be a QIVPM. For the classical case, if
$\ultramodular$ is not ultramodular, we also want to find a classical
probability measure $\mu:\events_{\mathrm{C}}\rightarrow\left[0,1\right]$ such
that $\ultramodular\circ\mu$ is not a classical IVPM. This can be
done by restricting our previous quantum probability measure
$\muB_{\rho}$ on a commuting subspace of
event~$\events_{\mathrm{C}}$ generated by $\proj{0}$, $\proj{1}$, $\proj{2}$,
and $\proj{3}$. The restricted function
$\mu=\muB_{\rho}|_{\events_{\mathrm{C}}}$ is a classical probability
measure, and the induced map $\ultramodular\circ\mu$ fails to be
a classical IVPM by the same reason as the quantum case.

>>>>

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\gerardo}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{OliveGreen}{Gerardo says: #1}\end{minipage}}}
\newcommand{\andy}[1]{\fbox{\begin{minipage}{0.9\linewidth}\color{blue}{Andy says: #1}\end{minipage}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{L{#1}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{R{#1}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{d}
\newcommand{\rmi}{i}
\lineskiplimit=1pt
\newcommand{\ultramodular}{\mathcal{M}}
\newcommand{\ultramodularL}[1][]{\ensuremath{\ultramodular^{L{#1}}}}
\newcommand{\ultramodularR}[1][]{\ensuremath{\ultramodular^{R{#1}}}}
\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\eventsC}{\ensuremath{\events_{C}}}

@@

We now have the necessary ingredients to define the quantum extension,
QIVPMs, as a generalization of both classical IVPMs and conventional
quantum probability measures. We will show that QIVPMs reduce to
classical IVPMs when the space of quantum events $\events$ is
restricted to mutually commuting events $\eventsC$, i.e.,
to compatible events that can be measured simultaneously. In
Sec.~\ref{sec:Gleason} we will discuss the connection between QIVPMs
and conventional quantum probability measures in detail.

@@

The proof that QIVPMs become classical IVPMs when the space of events
includes only mutually commuting events is presented in detail in the
long version of this paper~\cite{HOSTunpublished}. The idea of the
proof is as follows. Given a subspace of mutually commuting
events~$\eventsC$, we can construct a partial orthonormal
basis by diagonalization and complete it to a full orthonormal basis
$\overline{\eventsC}$. We then build a bijection between
the QIVPM on the set of projectors associated with this basis and the
set of classical events corresponding to this basis. This
correspondence leads to a crucial property in our setting.

\begin{thm}[Non-empty Core for Compatible Measurements] \label{thm:Shapley}
  For QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace
  of events~$\eventsC\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\eventsC\right)\ne\emptyset$.
\end{thm}

This theorem follows directly from our
correspondence~\cite{HOSTunpublished} and Shapley's
theorem~\cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997,Grabisch2016},
which proves that a classical IVPM always contains at least one state $\rho$
that is consistent with every event in $\eventsC$. In contrast, as we
illustrate in Sec.~\ref{sec:Gleason}, the set of states
$\coreBorn\left(\bar{\mu},\events'\right)$ that is consistent with
some arbitrary (not necessarily commuting) events $\events'$ and a
particular QIVPM $\bar{\mu}$ may be \emph{empty}.

@@

\begin{thm}[Equivalence of Ultramodular Functions and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}A function~$\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$
is ultramodular.
\item \label{enu:iota-mu-CIVPM}The composite
  function~$\ultramodular\circ\mu:\eventsC\rightarrow\mathscr{I}$
  is a classical IVPM for all classical probability
  measure~$\mu:\eventsC\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composite
  function~$\ultramodular\circ\mu:\events\rightarrow\mathscr{I}$ is a
  QIVPM for all quantum probability
  measure~$\mu:\events\rightarrow\left[0,1\right]$.
\end{enumerate}
\end{thm}

Statement~\ref{enu:iota-subject-to} implies \ref{enu:iota-mu-CIVPM}
and~\ref{enu:iota-mu-QIVPM} as we have outlined above. Conversely, for the quantum case, we want
to show that if $\ultramodular$ is not ultramodular, then for
some quantum probability measure $\mu$, the composite
$\ultramodular\circ\mu$ might not be a QIVPM. Suppose there are three
particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $y=x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but they don't
satisfy Eq.~(\ref{eq:iota-Inclusion}). Consider the state:
\[
\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-y\right)\proj{3}\,.
\]
The induced map using the Born rule and blurred by $\ultramodular$,
$\ultramodular\circ\muB_{\rho}$, fails to satisfy
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}) when
$P_{0}=\proj{0}+\proj{2}$ and $P_{1}=\proj{1}+\proj{2}$. In other
words, this induced map fails to be a QIVPM. For the classical case, if
$\ultramodular$ is not ultramodular, we also want to find a classical
probability measure $\mu:\eventsC\rightarrow\left[0,1\right]$ such
that $\ultramodular\circ\mu$ is not a classical IVPM. This can be
done by restricting our previous quantum probability measure
$\muB_{\rho}$ on a commuting subspace of
event~$\eventsC$ generated by $\proj{0}$, $\proj{1}$, $\proj{2}$,
and $\proj{3}$. The restricted function
$\mu=\muB_{\rho}|_{\eventsC}$ is a classical probability
measure, and the induced map $\ultramodular\circ\mu$ fails to be
a classical IVPM by the same reason as the quantum case.
